The \project{unRAVE} catalog

### 2016-09-05 article/unrave.tex
Write about gold standard comparisons

\begin{abstract}
The Milky Way is a powerful laboratory for understanding galaxy formation,
as it can provide orbits, ages, physical parameters, as well as chemical 
abundances for vast sets of individual stars.  These inferences require 
both astrometric and spectroscopic data.  Indeed, the Tycho-Gaia 
Astrometric Solution (\tgas) sample in the first \gaia\ data release 
requires a spectroscopic complement that includes radial velocities, 
stellar parameters, and elemental abundances.  Among existing spectroscopic 
samples, the RAdial Velocity Experiment (\rave) survey has the largest 
overlap with \tgas: 292,036 stars.  Here we present a data-driven 
re-analysis of \rave\ spectra, using an implementation of \thecannon, that
yields more precise and accurate stellar parameters and abundances.  We 
derive, and validate, improved effective temperature $\teff$, surface 
gravity $\logg$, and chemical abundances of up to seven elements (O, Mg, 
Al, Si, Ca, Fe, Ni) for \Nstarsqc\ stars.  The typical precision in chemical 
abundances is X.XX~dex for stars on the giant branch.  We provide 
spectrophotometric distances based on our stellar parameters, which should 
be superior to the parallax-based distances for faint or distant objects in
\tgas.  The \unrave\ catalog presented here, in conjunction with \tgas, will 
be the most powerful data set for chemo-orbital analysis at the dawn of 
broadly available \gaia\ data.
\end{abstract}

@@

\subsubsection{Comparison with \rave\ DR5}

Here we cross-match our results against literature samples to verify that
our results are accurate.  Because this work is an independent analysis, our
first point of reference is against the official \rave\ data release.  This
comparison is shown in Figure \ref{fig:rave-dr5-comparison}, 
where we show our results with respect to the fifth \rave\ data release \citep{Kunder_2016}.  In order
to provide a fair comparison, we only show the $XXX$ stars that meet a number of quality
flags in both samples: 
\stub{What selections qualify as a fair comparison in both samples?}
There is very good agreement in $\teff$, with a bias
and RMS of just XX~K and XXX~K, respectively.  We find our $\logg$ values are
closest to the `calibrated' \rave\ surface gravities, which make use of 
asteroseismic information to correct (increase) the $\logg$ values along the
giant branch.  The calibrated \rave\ surface gravities also correct the 
location of the red clump (by $\sim0.5$~dex) to be in the same position in our
results.  However, while our $\logg$ values generally agree with the calibrated
\rave\ values along the giant branch, our $\logg$ values do differ slightly 
along the main-sequence.  This is principally because the main-sequence in 
this work tends to taper down towards higher $\logg$ values at cooler 
temperatures, whereas the comparison sample tend to have a slightly flatter
main-sequence.  This difference is not likely to have a very significant 
effect on the detailed abundance or spectrophotometric distance determinations
between these studies \citep{Binney_2014}.

\subsubsection{Comparison with the \citet{Kordopatis_2013} calibration sample}
\label{sec:validation-kordopatis}

@@

\stub{Caution re MP dwarfs appear as sub-giants in our sample}

>>>>

\begin{abstract}
The Milky Way is a powerful laboratory for understanding galaxy formation,
as it can provide orbits, ages, physical parameters, as well as chemical 
abundances for vast sets of individual stars.  These inferences require 
both astrometric and spectroscopic data.  Indeed, in order to be maximally
useful for chemo-dynamic studies, the Tycho-Gaia Astrometric Solution 
(\tgas) sample in the first \gaia\ data release requires a spectroscopic 
complement that includes radial velocities, stellar parameters, and 
elemental abundances.  Among existing spectroscopic samples, the RAdial 
Velocity Experiment (\rave) survey has the largest overlap with \tgas: 
292,036 stars.  Here we present a data-driven re-analysis of \rave\ 
spectra, using an implementation of \thecannon, that yields more precise 
and accurate stellar parameters and abundances.  We derive, and validate,
improved effective temperature $\teff$, surface gravity $\logg$, and 
chemical abundances of up to seven elements (O, Mg, Al, Si, Ca, Fe, Ni) 
for \Nstarsqc\ stars.  The typical precision in chemical abundances is 
X.XX~dex for stars on the giant branch.  
The \unrave\ catalog presented here, in conjunction with \tgas, will 
be the most powerful data set for chemo-orbital analysis at the dawn of 
broadly available \gaia\ data.
\end{abstract}

%We *will* provide 
%spectrophotometric distances based on our stellar parameters, which should 
%be superior to the parallax-based distances for faint or distant objects in
%\tgas.

@@

\subsubsection{Comparison with \rave\ DR4}
\label{sec:validation-kordopatis}

As an initial point of comparison, we cross-matched our results against the 
official fourth \rave\ data release (Figure \ref{fig:rave-dr4-comparison}).
In order to provide a fair comparison, we only show stars that meet a number
of quality flags in \emph{both} samples. Specifically we require that the 
$S/N$ (as measured by our analysis) exceeds 10~pixel$^{-1}$, the reduced
$\chi_r^2$ in our analysis $< 3$, and that the \texttt{QK} flag reported by
\citet{Kordopatis_2013} is zero, indicating no problems with quality control.
There is very good agreement in $\teff$, with a bias and RMS of just XX~K 
and XXX~K, respectively.  The offset in $\logg$ on the giant branch between this study and 
\citet{Kordopatis_2013} has been noted in other studies, and this has issue
has been resolved in the fifth \rave\ data release by correcting $\logg$
values with a small sample of asteroseismic targets.  There is also a
slight discrepancy in the $\log$ values along the main-sequence, where our
work tends to taper down towards higher $\logg$ values at cooler temperatures,
and the \rave\ DR4 sample tends to have a slightly flatter lower main-sequence.
This difference is not likely to have a very significant effect on the detailed
abundance or spectrophotometric distance determinations between these studies 
\citep{Binney_2014}.

@@

We searched the literature for studies that overlap with \rave, and which base
their analysis on high-resolution, high S/N spectra.  We found four notable
studies with a sufficient level of overlap: the disk studies \citet{Reddy_2003}
and \citet{Reddy_2006}, the \citet{Valenti_Fischer_2005} work on exoplanet
host star candidates, and the work of \citet{Bensby_2014}.  These studies 
perform a careful (manual; expert) analysis using extremely high-resolution
\stub{how high?}, high S/N spectra, and make use of \project{Hipparcos}
parallaxes where possible.  Most of the stars in these samples are main-sequence
or sub-giant stars.  Therefore, these works constitute an excellent comparison
to evaluate the accuracy of our results on the main-sequence and sub-giant branch.

We compare the stellar parameters ($\teff$, $\logg$, [Fe/H]) of stars that
overlap these literature studies in Figure~\ref{fig:gold-standard-comparison}.
We only include stars with $\chi_r^2 < 3$ and $S/N > 10$, although the latter
cut removed only few stars because the average S/N in the \rave\ spectra for
these stars is relatively high ($\gtrsim{}50$). When considering the relative 
information content available in \rave\ (945~pixels in the near infrared with
$\mathcal{R} ~ 7{,}500$) compared to these literature studies (\stub{what snr, resolution, wavelength range and hipparcos parallaxes}, we consider the
 agreement to be satisfactory.

There are some discrepancies between our results and these literature studies,
and these differences arise from the manner in which we join results from the
main-sequence and giant model.  Specifically, for metal-poor stars near
the turn-off in these literature comparisons, in our weighted results we find 
these stars to have similar metallicities, but to be slightly more evolved: 
on the sub-giant branch or the base of the giant branch.  This effect is best
illustrated in Figure~\ref{fig:gold-standard-hrd}, where we show a 
Hertzsprung-Russell diagram for all stars in common between this work and that
of \citet{Bensby_2014,Reddy_2003,Reddy_2006,Valenti_Fischer_2005}.  While the
overall shape of the turn-off in our work is consistent with the literature
studies, the linked data points show that metal-poor stars in the literature
appear as more evolved stars in our combined results.  Instead, if we only
take the results from our main-sequence model --- ignoring the influence of the
giant model --- and compare those values to these literature studies, no such 
systematic discrepancies are present in the Hertzsprung-Russell diagram: we 
find those metal-poor turn-off stars to also be turn-off stars, with comparable
metallicities.  Our investigation indicates that this issue affects dwarf stars
near the turn-off that have $[{\rm Fe/H}] \lesssim -1$.  While the agreement
would improve if we adopted the main-sequence results for these specific stars,
the underlying issue of metal-poor dwarf stars would still be present.  Thus,
we use this example to illustrate caution for studies making inferences based
on main-sequence dwarf stars.  The results from both the main-sequence star model, 
and the giant star model, are available at \stub{should we put them online somewhere
for open-science, or just make them available upon request so that we don't
muddy the waters and confuse people as to which catalog to use?}
